Effects of diversity, topography, and interannual climate variability on pathogen spillover

### 2016-02-25 esa-2016-abstract.Rmd
Update results/conclusions.

## Results/Conclusions
Results from our path model of disease observations aggregated to the plot level revealed that diversity moderated pathogen spillover through a relatively strong direct negative effect on oak infection. Potential inoculum load had a direct positive effect on oak infection moderated by temperature and topography. Temperature and rainfall had relatively weaker influences compared to diversity and inoculum load in the context of pathogen spillover. Although diversity also had a positive effect on inoculum load, its negative effect on oak infection is consistent with the dilution effect found in other studies. Topography also had substantial influence on diversity and inoculum load, where higher values of the wetness index tended to have lower values of diversity, but higher inoculum load. This is consistent with areas where moisture is likely to accumulate and persist providing a more favorable environment for *P. ramorum* sporulation.

>>>>

## Results/Conclusions
We assessed our initial path model using measurements summarized for each plot. Our analysis indicated that diversity reduced pathogen spillover through a relatively strong direct negative effect on oak infection, however, this was tempered by more diverse plots also having a higher pathogen load. Pathogen load had the strongest direct positive effect on oak infection, which was countered by direct negative influences from temperature, and interestingly, bay laurel density. Our temperature variable was calcuated as the averaged daily maximum temperature during June-September preceding the spring sampling season, a time when bay laurel will drop damaged leaves. Our hypothesis was that higher temperatures leads to more leaf abscision and fewer infections persisting until the wet season. We tested the number of rainy days during November-May for each sampling season as our initial rainfall variable, because persistent moisture is more favorable for pathogen reproduction. However, this had a relatively weak effect on pathogen load. So far, we have confirmed a dilution effect via diversity reducing likelihood of oak infection, pathogen load has the single greatest effect on pathogen spillover, and discovered a surprising relationship between bay laurel and oak infection warranting further exploration.
